---
title: "Sensitivity analysis of MetaWards with an emulator"
author: Doug McNeall @dougmcneall dougmcneall@gmail.com
output: html_document
---

This code takes a small ensemble of runs of MetaWards and fits an emulator to the maximum number of infections in each run. The code then does a sensitivty analysis using the FASTT99 algorithm, and emulated output.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Packages   

```{r, message=FALSE, warning=FALSE}
library(sensitivity)
library(DiceKriging)
library(dplyr)
source("https://raw.githubusercontent.com/dougmcneall/packages-git/master/emtools.R")
```

### Read the design matrix from a designed ensemble of MetaWards, created in file "create_design_MetaWards.Rmd"
```{r}
# Need to fix the parameter names
X <- read.csv('design.csv', sep = "")
parnames = colnames(X)

```

### Read and summarise the results from the runs 
```{r}
# A container for all the data
# Each row has a "fingerprint" that contains the values of all the changed parameters,
# and the values of the parameters are also given.
# This alters the order of the parameters.
dat <- read.csv('output/results.csv.bz2')


# find an
unique_fingerprint = unique(dat$fingerprint)

# Use the dplyr package to find the maximum number of infections for each ensemble member.
max_infections <- dat %>% 
                      group_by(fingerprint) %>%
                      summarize(max(I))

reorder_ix <- match(unique_fingerprint, max_infections$fingerprint)
max_infections <- max_infections[reorder_ix, ]

head(max_infections)

```

Plot each parameter against the output to get an idea of sensitivity
```{r, fig.width = 10, fig.height = 8}

d <- ncol(X)

X.norm <- normalize(X)
y <- pull(max_infections,'max(I)')

par(mfrow = c(2,3), las = 1)

for(i in 1:d){
  plot(X[ ,i], y, xlab = parnames[i], ylab = 'max(I)')
}

```

### Fit the emulator
```{r, fig.width = 6, fig.height = 6, message=FALSE, warning=FALSE}

# Fit an emulator using DiceKriging
fit = km(~., design=X.norm, response=y)

# a quick check of the emulator using cross validation
loo = leaveOneOut.km(fit, type = 'UK', trend.reestim = TRUE)
```


```{r, fig.width = 6, fig.height = 6}
ylim = range(loo$mean - (2*loo$sd),loo$mean + (2*loo$sd) )
plot(y, loo$mean, xlab = 'max(I)', ylab = 'emulator prediction', ylim = ylim)
segments(x0 = y, y0 = loo$mean - (2*loo$sd), x1 = y, y1 = loo$mean + (2*loo$sd))
abline(0,1)

```

### FAST99 sensitivity analysis of Saltelli et al (1999)
```{r}

# Generate a design for the FAST99 analysis
X.fast <- fast99(model = NULL, factors = colnames(X), n = 1000,
                 q = "qunif", q.arg = list(min = 0, max = 1))


# Predict the response at the FAST99 design points using the emulator
pred.fast = predict(fit, newdata = X.fast$X, type = 'UK')

# Calculate the sensitivity indices
fast.tell <- tell(X.fast, pred.fast$mean)

bp.convert <- function(fastmodel){
  # get the FAST summary into an easier format for barplot
  fast.summ <- print(fastmodel)
  fast.diff <- fast.summ[ ,2] - fast.summ[ ,1]
  fast.bp <- t(cbind(fast.summ[ ,1], fast.diff))
  fast.bp
}

par(las = 2, mar = c(9,5,3,2))
barplot(bp.convert(fast.tell), col = c('skyblue', 'grey'), ylab = 'relative sensitivity')
legend('topleft',legend = c('Main effect', 'Interactions'), fill = c('skyblue', 'grey') )

```




