---
title: "MetaWards wave 03 pre-lockdown emulation test"
output: html_notebook
---


## Emulation tests for MetaWards

Model description (parameters etc. ) are [here](https://uq4covid.github.io/vignettes/data/MetaWards/vignette/ModelDescription)
Based on Code from James Salter

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(RSQLite)
library(fields)
library(viridisLite)
library(rgdal)
library(dplyr)
library(reshape2)
library(ggplot2)
library(cowplot)
library(DiceKriging)
library(emtools)
library(imptools)
library(viztools)

```


```{r}
#### Load in database ####
filename <- "data/wave03_summaries_DW.db" # likely just called wave_summaries_DW.db when downloaded, I've renamed
sqlite.driver <- dbDriver("SQLite")
db <- dbConnect(sqlite.driver,
                dbname = filename)
new_wave1_seeds <- dbReadTable(db,"compact")

# If instead have ward-level summaries
#filename <- "data/summaries_default.db"
#sqlite.driver <- dbDriver("SQLite")
#db <- dbConnect(sqlite.driver,
#                dbname = filename)
#dbListTables(db)

```

```{r, eval = FALSE}

newtable <- dbGetQuery(db, "SELECT * FROM compact WHERE week = 12 AND age = 8")

# Don't care about this column
new_wave1_seeds$Hprev_mn <- NULL

# Sum deaths
new_wave1_seeds$deaths <- new_wave1_seeds$cumHD + new_wave1_seeds$cumCD

#### Aggregate for each week ####
ExtractWeek <- function(all_data, weeks = c(12:6)){
  extracted_data <- NULL
  for (w in 1:length(weeks)){
    # First extract all those with replicates (NA else)
    tmp_data <- subset(all_data, week == weeks[w] & !(is.na(replicate)))
    tmp_data <- aggregate(cbind(cumH,cumHD,cumCD,deaths) ~ output + replicate, data = tmp_data, FUN = sum) # aggregate across age and trust
    tmp_data_reps <- aggregate(cbind(cumH,cumHD,cumCD,deaths) ~ output, data = tmp_data, FUN = median) # take median across replicates
    
    # Repeat for single runs (NAs)
    tmp_data <- subset(all_data, week == weeks[w] & (is.na(replicate)))
    tmp_data_unique <- aggregate(cbind(cumH,cumHD,cumCD,deaths) ~ output, data = tmp_data, FUN = sum) # aggregate across trusts
    
    # Combine
    tmp_data <- rbind(tmp_data_reps, tmp_data_unique)
    tmp_data$week <- weeks[w]
    
    extracted_data <- rbind(extracted_data, tmp_data)
  }
  
  return(extracted_data)
}

all_week_seeds <- ExtractWeek(new_wave1_seeds, weeks = c(12:6))
save(all_week_seeds, file = 'data/all_week_seeds.RData')


#### Aggregate final week by age ####
ExtractAge <- function(all_data){
  # First extract all those with replicates (NA else)
  tmp_data <- subset(all_data, week == 12 & !(is.na(replicate)))
  tmp_data <- aggregate(cbind(cumH,cumHD,cumCD,deaths) ~ output + age + replicate, data = tmp_data, FUN = sum) # aggregate across trust
  tmp_data_reps <- aggregate(cbind(cumH,cumHD,cumCD,deaths) ~ output + age, data = tmp_data, FUN = median) # take median across replicates   
  # Repeat for single runs (NAs)
  tmp_data <- subset(all_data, week == 12 & (is.na(replicate)))
  tmp_data_unique <- aggregate(cbind(cumH,cumHD,cumCD,deaths) ~ output + age, data = tmp_data, FUN = sum) # aggregate across trusts
  
  # Combine
  tmp_data <- rbind(tmp_data_reps, tmp_data_unique)
  return(tmp_data)
}

week12age_seeds <- ExtractAge(new_wave1_seeds)
save(week12age_seeds, file = 'data/week12age_seeds.RData')
```

```{r}


load('data/all_week_seeds.RData') # dataframe of weekly cumulative totals
load('data/week12age_seeds.RData') # dataframe of week 12 totals by age
all_week12 <- subset(all_week_seeds, week == 12) # week 12 cumulative totals only

#### Load in design ####
NewDesign <- read.csv('data/wave03inputs.csv') # again, might be called something else

# Merge design and model output
CombinedData <- all_week_seeds %>% left_join(NewDesign, by = c('output' = 'output'))

# Load csv file of LAD deaths, trust admissions (not public)
#obs_deaths <- read.csv('data/forTJ-lad-deaths-by-age-2021-03-22.csv') # consistent, depending on 28 day vs certificate definition
#trust_admissions <- read.csv('data/forTJ-trust-admissions-by-age-2021-03-22.csv')

# Download various PHE datasets
phe_deaths <- read.csv('https://api.coronavirus.data.gov.uk/v2/data?areaType=nation&metric=cumDeaths28DaysByDeathDate&metric=newDeaths28DaysByDeathDate&format=csv')
phe_deaths <- subset(phe_deaths, areaName == 'England') # probbly want Wales as well really

phe_deaths_cert <- read.csv('https://api.coronavirus.data.gov.uk/v2/data?areaType=nation&metric=newDailyNsoDeathsByDeathDate&metric=cumDailyNsoDeathsByDeathDate&format=csv')
phe_deaths_cert <- subset(phe_deaths_cert, areaName == 'England') # again, need Wales as well when this isn't such a small number

phe_data <- read.csv('https://api.coronavirus.data.gov.uk/v2/data?areaType=nation&metric=newAdmissions&metric=cumAdmissions&format=csv')
phe_data <- subset(phe_data, areaName == 'England')

phe_admissions <- read.csv('https://api.coronavirus.data.gov.uk/v2/data?areaType=nation&areaCode=E92000001&metric=newAdmissions&format=csv')

```



```{r, eval = FALSE}
#### Subset all datasets so matches MW output ####
EndDate <- as.Date('2020-03-21')

#trust_admissions <- subset(trust_admissions, date <= EndDate)
#trust_admissions <- aggregate(value ~ date, trust_admissions, sum) # summing across trusts

# byDayDeaths <- aggregate(value ~ date, data = obs_deaths, sum, na.rm=TRUE) # summing across LADs

# Extract weekly numbers from observations, on dates that match with MW (week 12 is shorter)
date_seq <- c(seq(from = as.Date('2020-03-17'), by = -7, len = 6)[6:1], EndDate)
totalDeaths <- numeric(7)
for (i in 1:7){
  totalDeaths[i] <- sum(subset(byDayDeaths, as.character(date) <= date_seq[i])$value)
}

# The PHE datasets start at different times, so this is slightly messy and not general - be careful with zeros!!!
obs_data <- data.frame(week = 6:12, 
                       median = log(subset(phe_deaths_cert, as.Date(date) %in% date_seq)$cumDailyNsoDeathsByDeathDate[7:1]+1), 
                       lower = c(0,0,0,0,log(subset(phe_deaths, as.Date(date) %in% date_seq)$cumDeaths28DaysByDeathDate[3:1]+1)), 
                       upper = log(totalDeaths+1))
```


```{r, eval = FALSE, fig.height=6, out.width='50%', fig.show='hold'}
ggplot(subset(CombinedData, week == 12), aes(x = log(cumH + 1), y = log(deaths + 1), col = ns)) +
  geom_point() +
  geom_hline(yintercept = c(log(250+1), log(2500+1))) +
  geom_vline(xintercept = c(log(2500+1), log(10000+1))) +
  scale_color_viridis() +
  labs(col = 'Seeds', x = 'log(hosps)', y = 'log(deaths)')

ggplot(subset(CombinedData, week == 12), aes(x = log(cumH + 1), y = log(deaths + 1), col = R0)) +
  geom_point() +
  geom_hline(yintercept = c(log(250+1), log(2500+1))) +
  geom_vline(xintercept = c(log(2500+1), log(10000+1))) +
  scale_color_viridis() +
  labs(x = 'log(hosps)', y = 'log(deaths)')
```


### Cumulative hospital deaths
```{r, fig.width = 12, fig.height = 8}

week12 <- subset(CombinedData, week == 12)

par(mfrow = c(3,5))

for (i in 1:14 ){
  plot(NewDesign[,i], week12$cumHD, xlab = colnames(NewDesign)[i], ylab = 'cumulative hospital deaths')
}

```



### Cumulative community deaths
```{r, fig.width = 12, fig.height = 8}

par(mfrow = c(3,5))

for (i in 1:14 ){
  plot(NewDesign[,i], week12$cumCD, xlab = colnames(NewDesign)[i], ylab = 'cumulative community deaths')
}

```

```{r}



```

## Emulating cumulative deaths in hostpital (cumHD) in the community (cumCD) at the beginning of lockdown
The DiceKriging Emulator fit looks good
```{r, fig.width = 5, fig.height = 8,  include = FALSE}

fit_cumHD <- km(~., design = NewDesign[, 1:14], response = log(1+ week12$cumHD))

fit_cumCD <- km(~., design = NewDesign[, 1:14], response = log(1+ week12$cumCD))

```

```{r}
loo_cumHD <- leaveOneOut.km(fit_cumHD, type = 'UK', trend.reestim = TRUE)
loo_cumCD <- leaveOneOut.km(fit_cumCD, type = 'UK', trend.reestim = TRUE)

```

### Week 12 cumulative Hospital deaths Leave-one-out plots
```{r, fig.width = 12, fig.height = 6}


par(mfrow = c(1,2))
plot(log(1+ week12$cumHD), loo_cumHD$mean, main = 'LOO log(1+ Cumulative Hospital Deaths)')
segments(x0 =log(1+ week12$cumHD), y0 = loo_cumHD$mean - (2* loo_cumHD$sd), 
        x1 =log(1+ week12$cumHD), y1 = loo_cumHD$mean + (2* loo_cumHD$sd)
)
abline(0,1)


y0 = exp(loo_cumHD$mean) - (exp(loo_cumHD$mean) * (exp(loo_cumHD$sd) - 1))
y1 = exp(loo_cumHD$mean) + (exp(loo_cumHD$mean) * (exp(loo_cumHD$sd) - 1))

plot(week12$cumHD, exp(loo_cumHD$mean), main = 'LOO Cumulative Hospital Deaths', ylim = range(y0, y1) 
     )
segments(x0 = week12$cumHD, y0 = y0,
         x1 = week12$cumHD, y1 = y1
)

abline(0,1)


```


### Week 12 cumulative community deaths Leave-one-out plots
```{r, fig.width = 12, fig.height = 6}

par(mfrow = c(1,2))
plot(log(1+ week12$cumCD), loo_cumCD$mean, main = 'LOO log(1+ Cumulative community Deaths)')
segments(x0 =log(1+ week12$cumCD), y0 = loo_cumCD$mean - (2* loo_cumCD$sd), 
        x1 =log(1+ week12$cumCD), y1 = loo_cumCD$mean + (2* loo_cumCD$sd)
)
abline(0,1)


y0 = exp(loo_cumCD$mean) - (exp(loo_cumCD$mean) * (exp(loo_cumCD$sd) - 1))
y1 = exp(loo_cumCD$mean) + (exp(loo_cumCD$mean) * (exp(loo_cumCD$sd) - 1))

plot(week12$cumCD, exp(loo_cumCD$mean), main = 'LOO Cumulative community Deaths', ylim = range(y0, y1) 
     )
segments(x0 = week12$cumCD, y0 = y0,
         x1 = week12$cumCD, y1 = y1
)

abline(0,1)


```

### Ensemble design

```{r, fig.width = 12, fig.height = 12}
pairs(NewDesign[,1:14])

```

```{r}

X <- NewDesign[, 1:14]
X_norm <- normalize(X)

fit_cumHD_norm <- km(~., design = X_norm, response = log(1+ week12$cumHD))

fit_cumCD_norm <- km(~., design = X_norm, response = log(1+ week12$cumCD))


```

```{r, fig.width = 12, fig.height = 12}


X_oaat <- oaat_design(X_norm, n=21, med = TRUE)
colnames(X_oaat) = colnames(X)

y_oaat <- predict.km(fit_cumHD_norm, newdata = X_oaat, type = 'UK')

oaatLinePlot(X_oaat = X_oaat, y_oaat_mean = y_oaat$mean, y_oaat_sd = y_oaat$sd, 
             n_oaat = 21,nr = 3, nc = 5) 
```


```{r, fig.width = 12, fig.height = 12}

y_oaat <- predict.km(fit_cumCD_norm, newdata = X_oaat, type = 'UK')

oaatLinePlot(X_oaat = X_oaat, y_oaat_mean = y_oaat$mean, y_oaat_sd = y_oaat$sd, 
             n_oaat = 21,nr = 3, nc = 5) 
```


