---
title: "Local Coronavirus Tracker"
author: "Doug McNeall"
contact: "dougmcneall@gmail.com @dougmcneall"
date: "2021-04-28               "
output: html_document   
---
        
## Introduction
Confirmed cases of COVID-19 in England. Note these are lab confirmed cases only - the real number of cases will be higher, and the numbers for a recent day could be revised upwards at a later date. In particular, the "about the data" link suggests that data older than 5 days can be considered complete.
                
Data from <https://coronavirus.data.gov.uk/>  
About the data <https://coronavirus.data.gov.uk/about>  
Data file <https://coronavirus.data.gov.uk/downloads/csv/coronavirus-cases_latest.csv>  
Code to create this page <https://github.com/dougmcneall/covid/blob/master/docs/local_coronavirus.Rmd>

This data will be used to calibrate [MetaWards](https://metawards.org/), an infectious disease model being used to understand and predict COVID-19. You can see progress on our uncertainty quantification project using MetaWards [here](https://uq4covid.github.io/).

Contact: dougmcneall@gmail.com

### Update 2021-01-18
Lines now show 7-day rolling means, with the last week excluded as before.

### Update 2020-07-29
Weekly averages are now included. The last week's average data is not plotted, as it is often incomplete giving a misleading drop in cases.
        
### Update 2020-07-03
Data and graphs now include all cases from both pillar 1 (hospital settings) and pillar 2 (private labs and home testing), backdated through the record. This may well have increased the number of recorded cases in your local area considerably.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(lubridate)
library(zoo)
```


```{r, include=FALSE}
#dat <- read.csv('https://coronavirus.data.gov.uk/downloads/csv/coronavirus-cases_latest.csv')

#dat <- read.csv('https://coronavirus.data.gov.uk/downloads/csv/coronavirus-cases_latest.csv')

# Having to download a copy of the csv from the web adress above because it's broken, for some reason.
download.file('https://coronavirus.data.gov.uk/downloads/csv/coronavirus-cases_latest.csv', destfile='coronavirus-cases_latest.csv')
dat <- read.csv('coronavirus-cases_latest.csv')

#weekly <- read.csv('../data/weekly_uk_covid19_cases_2020-07-02.csv')
```


```{r, include=FALSE}

plotLocalCases <- function(area = 'Devon', dat, col = 'skyblue2', ...){
        # 'area' should be a string that names an area of interest
        
        localDataAll <- dat[dat[,'Area.name']==area, ]
        
        # The Unitary Authorities (eg   'Dorset', 'Plymouth') have the data doubled, and Area.type recorded as both 
        # 'Lower tier local authority' and 'Upper tier local Authority'. Remove one of them.
        
        authvec <- unique(localDataAll$Area.type)
        localData <- localDataAll[localDataAll$Area.type==authvec[1], ]
        
        da = seq(from = min(as.Date(localData[, 'Specimen.date'])), to = max(as.Date(localData[, 'Specimen.date'])), by = 7)

        par(las = 1, mar = c(2.5,4,5,1))
        plot(as.Date(localData[, 'Specimen.date']), localData[, 'Daily.lab.confirmed.cases'],
             xlab = '', ylab = 'cases', main = '',
             type = 'h', lwd = 5, col = col, bty = 'n', axes = FALSE, ...)
        grid()
        
        axis.Date(1,at=da,labels=format(da,"%d %b"),las=1)
        axis(2)
        mtext(text = paste0('Daily lab-confirmed Covid-19 cases in ', area), side = 3, line = .5, adj = 0, cex = 1.5)
}

```

```{r, include = FALSE}

plotLocalCasesWeeklyOverlay <- function(area = 'Devon', dat, col = 'skyblue2', ...){
        # 'area' should be a string that names an area of interest
        
        localDataAll <- dat[dat[,'Area.name']==area, ]
        
        # The Unitary Authorities (eg 'Dorset', 'Plymouth') have the data doubled, and Area.type recorded as both 
        # 'Lower tier local authority' and 'Upper tier local Authority'. Remove one of them.
        
        authvec <- unique(localDataAll$Area.type)
        localData <- localDataAll[localDataAll$Area.type==authvec[1], ]
        
        da = seq(from = min(as.Date(localData[, 'Specimen.date'])), to = max(as.Date(localData[, 'Specimen.date'])), by = 7)
        
        # plot daily data
        par(las = 1, mar = c(2.5,4,5,1))
        plot(as.Date(localData[, 'Specimen.date']), localData[, 'Daily.lab.confirmed.cases'],
             xlab = '', ylab = 'cases', main = '',
             type = 'h', lwd = 5, col = col, bty = 'n', axes = FALSE, ...)
        grid()
        
        axis.Date(1,at=da,labels=format(da,"%d %b"),las=1)
        axis(2)
        mtext(text = paste0('Daily lab-confirmed Covid-19 cases in ', area), side = 3, line = .5, adj = 0, cex = 1.5)
        
        # Summarise weekly data
        init <- data.frame('Specimen.date' = as.Date(localData[,'Specimen.date']), 
                           'Daily.lab.confirmed.cases' = localData[,'Daily.lab.confirmed.cases'])
        
        weekly_new_cases <- init %>% 
                group_by(week = week(Specimen.date)) %>%
                summarise(weekly_sum = sum(Daily.lab.confirmed.cases))
        
        
        week_ax <- as.Date(paste(2020, weekly_new_cases$week, 1, sep="-"), "%Y-%U-%u")
        
        #week_ax = seq(from = min(as.Date(localData[, 'Specimen.date'])), by = 7, length.out = nrow(weekly_new_cases))
        
        # plot but exclude the latest week using head( , -1)
        points(head(week_ax,-1), head(weekly_new_cases$weekly_sum/7,-1), pch = 20, col = 'skyblue4', type = 'o')
        legend('topright', legend = 'Weekly Average', pch = 20, col = 'skyblue4', lty = 'solid', bty = 'n')
        
}
```


```{r, echo=FALSE, include =FALSE, fig.width=12, fig.height = 4 }

# Hard code the double plot here for the minute
# isolate Exeter cases
exe <- dat[dat[,'Area.name']=='Exeter',]
devon <- dat[dat[,'Area.name']=='Devon',]

# Summarise weekly data
init_devon <- data.frame('Specimen.date' = as.Date(devon[,'Specimen.date']), 
                         'Daily.lab.confirmed.cases' = devon[,'Daily.lab.confirmed.cases'])


weekly_new_cases_devon <- init_devon %>% 
        group_by(week = week(Specimen.date)) %>%
        summarise(weekly_sum = sum(Daily.lab.confirmed.cases))

# Summarise weekly data
init_exe <- data.frame('Specimen.date' = as.Date(exe[,'Specimen.date']), 
                         'Daily.lab.confirmed.cases' = exe[,'Daily.lab.confirmed.cases'])

weekly_new_cases_exe <- init_exe %>% 
        group_by(week = week(Specimen.date)) %>%
        summarise(weekly_sum = sum(Daily.lab.confirmed.cases))

par(las = 1, mar = c(2.5,4,5,1))
plot(as.Date(devon[, 'Specimen.date']), devon[, 'Daily.lab.confirmed.cases'],
     xlab = '', ylab = 'cases', main = '',
     type = 'h', lwd = 5, col = 'skyblue2', bty = 'n', axes = FALSE)
grid()

da = seq(from = min(as.Date(devon[, 'Specimen.date'])), to = max(as.Date(devon[, 'Specimen.date'])), by = 7)
axis.Date(1,at=da,labels=format(da,"%d %b"),las=1)
axis(2)

points(as.Date(exe[, 'Specimen.date']), exe[, 'Daily.lab.confirmed.cases'],
     xlab = 'Date', ylab = 'cases', main = '',
     type = 'h', col = 'tomato2', lwd = 5)
legend('topleft', legend = c('Devon', 'Exeter'), col = c('skyblue2', 'tomato2'), lty = 'solid', lwd=5, bty = 'n')
mtext(text = 'Daily lab-confirmed Covid-19 cases in Devon & Exeter', side = 3, line = .5, adj = 0, cex = 1.5)

week_ax <- as.Date(paste(2020, weekly_new_cases_devon$week, 1, sep="-"), "%Y-%U-%u")

points(head(week_ax,-1), head(weekly_new_cases_devon$weekly_sum/7,-1), pch = 20, col = 'skyblue4', type = 'o')

week_ax <- as.Date(paste(2020, weekly_new_cases_exe$week, 1, sep="-"), "%Y-%U-%u")

points(head(week_ax,-1), head(weekly_new_cases_exe$weekly_sum/7,-1), pch = 20, col = 'tomato4', type = 'o')
legend('topright', legend = 'Weekly Average', pch = 20, col = 'skyblue4', lty = 'solid', bty = 'n')
        
```


```{r, echo = FALSE}
plotLocalCasesRollingOverlay <- function(area = 'Devon', dat, col = 'skyblue2', ...){
              # 'area' should be a string that names an area of interest
        
        localDataAll <- dat[dat[,'Area.name']==area, ]
        
        # The Unitary Authorities (eg 'Dorset', 'Plymouth') have the data doubled, and Area.type recorded as both 
        # 'Lower tier local authority' and 'Upper tier local Authority'. Remove one of them.
        
        authvec <- unique(localDataAll$Area.type)
        localData <- localDataAll[localDataAll$Area.type==authvec[1], ]
        
        localDataOrdered <-   localData[ order(as.Date(localData[,'Specimen.date'])),  ]
        
        LocalDataRollMean <- rollmean(localDataOrdered[, 'Daily.lab.confirmed.cases'], 7, align = 'center', na.pad = T)
        LocalDataRollMean[(length(LocalDataRollMean)-7):length(LocalDataRollMean)] <- NA
        
        da = seq(from = min(as.Date(localData[, 'Specimen.date'])), to = max(as.Date(localData[, 'Specimen.date'])), by = 7)
        
        # plot daily data
        par(las = 1, mar = c(2.5,4,5,1))
        plot(as.Date(localDataOrdered[, 'Specimen.date']), localDataOrdered[, 'Daily.lab.confirmed.cases'],
             xlab = '', ylab = 'cases', main = '',
             type = 'h', lwd = 5, col = col, bty = 'n', axes = FALSE, ...)
        grid()
        
        axis.Date(1,at=da,labels=format(da,"%d %b"),las=1)
        axis(2)
        mtext(text = paste0('Daily lab-confirmed Covid-19 cases in ', area), side = 3, line = .5, adj = 0, cex = 1.5)
        
        lines(as.Date(localDataOrdered[, 'Specimen.date']), LocalDataRollMean, col = 'skyblue4', lwd = 2)
        legend('topright', legend = '7 day rolling Average', pch = 20, col = 'skyblue4', lty = 'solid', bty = 'n')  
        
}

```

## The South West
```{r, fig.width=12, fig.height = 4, echo = FALSE}

       exe <- dat[dat[,'Area.name']=='Exeter',]
        exe.order <- exe[ order(as.Date(exe[,'Specimen.date'])),  ]
        
        exe.rm <- rollmean(exe.order[, 'Daily.lab.confirmed.cases'], 7, align = 'center', na.pad = T)
        # replace last 7 days with NA
        exe.rm[(length(exe.rm)-7):length(exe.rm)] <- NA
        
       # , format="%d/%m/%Y")
        
        devon <- dat[dat[,'Area.name']=='Devon',]
        devon.order <- devon[ order(as.Date(devon[,'Specimen.date'])),  ]
        
        devon.rm <- rollmean(devon.order[, 'Daily.lab.confirmed.cases'], 7, align = 'center', na.pad = T)
        devon.rm[(length(devon.rm)-7):length(devon.rm)] <- NA
        
        par(las = 1, mar = c(2.5,4,5,1))
        
        plot(as.Date(devon.order[, 'Specimen.date']), devon.order[, 'Daily.lab.confirmed.cases'],
           xlab = '', ylab = 'cases', main = '',
          type = 'h', lwd = 5, col = 'skyblue2', bty = 'n', axes = FALSE)
        grid()
        
        lines(as.Date(devon.order[, 'Specimen.date']), devon.rm, col = 'skyblue4', lwd = 2)
        
        points(as.Date(exe.order[, 'Specimen.date']), exe.order[, 'Daily.lab.confirmed.cases'],
     xlab = 'Date', ylab = 'cases', main = '',
     type = 'h', col = 'tomato2', lwd = 5)
        
        lines(as.Date(exe.order[, 'Specimen.date']), exe.rm, col = 'tomato4', lwd = 2)
        
        
legend('topleft', legend = c('Devon', 'Exeter'), col = c('skyblue2', 'tomato2'), lty = 'solid', lwd=5, bty = 'n')
mtext(text = 'Daily lab-confirmed Covid-19 cases in Devon & Exeter', side = 3, line = .5, adj = 0, cex = 1.5)
        
        
        da = seq(from = min(as.Date(devon.order[, 'Specimen.date'])), to = max(as.Date(devon.order[, 'Specimen.date'])), by = 7)
axis.Date(1,at=da,labels=format(da,"%d %b"),las=1)
axis(2)
 legend('topright', legend = '7 day rolling Average', pch = 20, col = 'skyblue4', lty = 'solid', bty = 'n')  
        
```



```{r, echo = FALSE, include = FALSE, fig.width = 12, fig.height = 4}
plotlist_sw = list('Dorset', 'Cornwall and Isles of Scilly', 'Somerset', 'Plymouth', 'Torbay','Bournemouth, Christchurch and Poole' )

out = lapply(plotlist_sw, FUN = plotLocalCasesWeeklyOverlay, dat = dat)

```


```{r, echo = FALSE, fig.width = 12, fig.height = 4}
plotlist_sw = list('Dorset', 'Cornwall and Isles of Scilly', 'Somerset', 'Plymouth', 'Torbay','Bournemouth, Christchurch and Poole' )

out = lapply(plotlist_sw, FUN = plotLocalCasesRollingOverlay, dat = dat)

```


## Other places of interest

```{r, echo=FALSE, fig.width=12, fig.height = 4 }

plotlist_opoi = list('Milton Keynes','Reading')

out = lapply(plotlist_opoi, FUN = plotLocalCasesRollingOverlay, dat = dat)


```

## England and English regions

```{r, echo=FALSE, fig.width=12, fig.height = 4 }

plotlist_eaer = list('England', 'South West', 'South East','London', 'East of England', 'West Midlands', 'East Midlands',
                     'Yorkshire and The Humber','North West', 'North East')

out = lapply(plotlist_eaer, FUN = plotLocalCasesRollingOverlay, dat = dat)

```


```{r, include=FALSE}

# testing statistics    
#tests <- read.csv(file = 'https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/906109/2020-08-02-COVID-19-UK-testing-time-series.csv')


#EnglandTests <- tests[tests[,'Nation']=='England', ]

#pillar2EnglandTests <- tests[tests[,'Nation']=='England' & tests[,'Pillar']=='Pillar 1', ]

```


```{r, include=FALSE}

plotLocalRates <- function(area = 'Devon', dat, col = 'skyblue2', ...){
        # 'area' should be a string that names an area of interest
        
        localData <- dat[dat[,'Area.name']==area, ]
        
        da = seq(from = min(as.Date(localData[, 'Specimen.date'])), to = max(as.Date(localData[, 'Specimen.date'])), by = 7)

        
        par(las = 1, mar = c(2.5,4,4,1))
        plot(as.Date(localData[, 'Specimen.date']), localData[, 'Cumulative.lab.confirmed.cases.rate'],
             xlab = '', ylab = 'cases per 100 000 population', main = paste0('Cumulative lab confirmed cases rate in ', area),
             type = 'h', lwd = 5, col = col, bty = 'n', axes = FALSE, ...)
        grid()
        
        axis.Date(1,at=da,labels=format(da,"%d %b"),las=1)
        axis(2)
}

plotLocalRates(area = 'Devon', dat = dat)

```


```{r, fig.width=10, fig.height=4, include=FALSE, echo=FALSE}

par(las = 1)
plot(as.Date(devon[, 'Specimen.date']), devon[, 'Daily.lab.confirmed.cases'], type = 'n', axes = FALSE,
     xlab = 'Date', ylab = 'Daily lab confirmed cases', main = 'Daily lab confirmed Coronavirus cases in Devon & Exeter')
grid()
axis.Date(1, devon[, 'Specimen.date'])
axis(2)
lines(as.Date(devon[, 'Specimen.date']), devon[, 'Daily.lab.confirmed.cases'], type = 'o', pch = 20, col = 'black', lwd = 1.5)
lines(as.Date(exe[, 'Specimen.date']), exe[, 'Daily.lab.confirmed.cases'], type = 'o', pch = 20, col = 'blue', lwd = 1.5)
legend('topleft', legend = c('Devon', 'Exeter'), col = c('black', 'blue'), lty = 'solid', lwd = 1.5, pch = 20)
```



