---
title: "Local Coronavirus Tracker"
author: "Doug McNeall"
contact: "dougmcneall@gmail.com @dougmcneall"
date: "2020-05-13"
output: html_document
---



Note these are lab confirmed cases only - the real number of cases will be higher, and the numbers for a recent day could be revised upwards at a later date.  

Data from <https://coronavirus.data.gov.uk/>  
About the data <https://coronavirus.data.gov.uk/about>  
Data file <https://coronavirus.data.gov.uk/downloads/csv/coronavirus-cases_latest.csv'>    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
dat <- read.csv('https://coronavirus.data.gov.uk/downloads/csv/coronavirus-cases_latest.csv')

# isolate Exeter cases
exe <- dat[dat[,1]=='Exeter',]
devon <- dat[dat[,1]=='Devon',]
dorset <- dat[dat[,1]=='Dorset',]

mk <- dat[dat[,1]=='Milton Keynes',]
hk <- dat[dat[,1]=='Hackney',]

```

### New daily lab-confirmed Coronavirus cases in Devon and Exeter  
```{r, echo=FALSE, fig.width=12, fig.height = 4 }
par(las = 1, mar = c(2.5,4,4,1))
plot(as.Date(devon[, 'Specimen.date']), devon[, 'Daily.lab.confirmed.cases'],
     xlab = '', ylab = 'cases', main = 'Daily lab-confirmed Coronavirus cases in Devon & Exeter',
     type = 'h', lwd = 5, col = 'skyblue2', bty = 'n')
grid()

points(as.Date(exe[, 'Specimen.date']), exe[, 'Daily.lab.confirmed.cases'],
     xlab = 'Date', ylab = 'cases', main = 'Daily lab confirmed Coronavirus cases in Devon & Exeter',
     type = 'h', col = 'tomato2', lwd = 5)
legend('topleft', legend = c('Devon', 'Exeter'), col = c('skyblue2', 'tomato2'), lty = 'solid', lwd=5, bty = 'n')

```


### New daily lab-confirmed Coronavirus cases in Dorset
```{r, echo=FALSE, fig.width=12, fig.height = 4 }
par(las = 1, mar = c(2.5,4,4,1))
plot(as.Date(dorset[, 'Specimen.date']), dorset[, 'Daily.lab.confirmed.cases'],
     xlab = '', ylab = 'cases', main = 'Daily lab-confirmed Coronavirus cases in Dorset',
     type = 'h', lwd = 5, col = 'skyblue2', bty = 'n')
grid()

```


### New daily lab-confirmed Coronavirus cases in Milton Keynes 
```{r, echo=FALSE, fig.width=12, fig.height = 4 }
par(las = 1, mar = c(2.5,4,4,1))
plot(as.Date(mk[, 'Specimen.date']), mk[, 'Daily.lab.confirmed.cases'],
     xlab = '', ylab = 'cases', main = 'Daily lab-confirmed Coronavirus cases in Milton Keynes',
     type = 'h', lwd = 5, col = 'skyblue2', bty = 'n')
grid()

```


### New daily lab-confirmed Coronavirus cases in Hackney
```{r, echo=FALSE, fig.width=12, fig.height = 4 }
par(las = 1, mar = c(2.5,4,4,1))
plot(as.Date(hk[, 'Specimen.date']), hk[, 'Daily.lab.confirmed.cases'],
     xlab = '', ylab = 'cases', main = 'Daily lab-confirmed Coronavirus cases in Hackney',
     type = 'h', lwd = 5, col = 'skyblue2', bty = 'n')
grid()

```


```{r, fig.width=10, fig.height=4, include=FALSE, echo=FALSE}

par(las = 1)
plot(as.Date(devon[, 'Specimen.date']), devon[, 'Daily.lab.confirmed.cases'], type = 'n', axes = FALSE,
     xlab = 'Date', ylab = 'Daily lab confirmed cases', main = 'Daily lab confirmed Coronavirus cases in Devon & Exeter')
grid()
axis.Date(1, devon[, 'Specimen.date'])
axis(2)
lines(as.Date(devon[, 'Specimen.date']), devon[, 'Daily.lab.confirmed.cases'], type = 'o', pch = 20, col = 'black', lwd = 1.5)
lines(as.Date(exe[, 'Specimen.date']), exe[, 'Daily.lab.confirmed.cases'], type = 'o', pch = 20, col = 'blue', lwd = 1.5)
legend('topleft', legend = c('Devon', 'Exeter'), col = c('black', 'blue'), lty = 'solid', lwd = 1.5, pch = 20)
```



