---
title: "Local Coronavirus Tracker"
author: "Doug McNeall"
contact: "dougmcneall@gmail.com @dougmcneall"
date: "2020-06-24"      
output: html_document   
---

Note these are lab confirmed cases only - the real number of cases will be higher, and the numbers for a recent day could be revised upwards at a later date. In particular, the "about the data" link suggests that data older than 5 days can be considered complete.

Data from <https://coronavirus.data.gov.uk/>  
About the data <https://coronavirus.data.gov.uk/about>  
Data file <https://coronavirus.data.gov.uk/downloads/csv/coronavirus-cases_latest.csv>  
Code to create this page <https://github.com/dougmcneall/covid/blob/master/docs/local_coronavirus.Rmd>

Contact: dougmcneall@gmail.com

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
dat <- read.csv('https://coronavirus.data.gov.uk/downloads/csv/coronavirus-cases_latest.csv')

```


```{r, include=FALSE}

plotLocalCases <- function(area = 'Devon', dat, col = 'skyblue2', ...){
        # 'area' should be a string that names an area of interest
        
        localData <- dat[dat[,'Area.name']==area, ]
        
        da = seq(from = min(as.Date(localData[, 'Specimen.date'])), to = max(as.Date(localData[, 'Specimen.date'])), by = 7)

        
        par(las = 1, mar = c(2.5,4,4,1))
        plot(as.Date(localData[, 'Specimen.date']), localData[, 'Daily.lab.confirmed.cases'],
             xlab = '', ylab = 'cases', main = paste0('Daily lab-confirmed Coronavirus cases in ', area),
             type = 'h', lwd = 5, col = col, bty = 'n', axes = FALSE, ...)
        grid()
        
        axis.Date(1,at=da,labels=format(da,"%d %b"),las=1)
        axis(2)
}

```


## Devon and Exeter  
```{r, echo=FALSE, fig.width=12, fig.height = 4 }

# Hard code the double plot here for the minute
# isolate Exeter cases
exe <- dat[dat[,'Area.name']=='Exeter',]
devon <- dat[dat[,'Area.name']=='Devon',]

par(las = 1, mar = c(2.5,4,4,1))
plot(as.Date(devon[, 'Specimen.date']), devon[, 'Daily.lab.confirmed.cases'],
     xlab = '', ylab = 'cases', main = 'Daily lab-confirmed Coronavirus cases in Devon & Exeter',
     type = 'h', lwd = 5, col = 'skyblue2', bty = 'n', axes = FALSE)
grid()


da = seq(from = min(as.Date(devon[, 'Specimen.date'])), to = max(as.Date(devon[, 'Specimen.date'])), by = 7)
axis.Date(1,at=da,labels=format(da,"%d %b"),las=1)
axis(2)

points(as.Date(exe[, 'Specimen.date']), exe[, 'Daily.lab.confirmed.cases'],
     xlab = 'Date', ylab = 'cases', main = 'Daily lab confirmed Coronavirus cases in Devon & Exeter',
     type = 'h', col = 'tomato2', lwd = 5)
legend('topleft', legend = c('Devon', 'Exeter'), col = c('skyblue2', 'tomato2'), lty = 'solid', lwd=5, bty = 'n')

```


## Dorset
```{r, echo=FALSE, fig.width=12, fig.height = 4 }

plotLocalCases(area = 'Dorset', dat = dat)

```


## South West
```{r, echo=FALSE, fig.width=12, fig.height = 4 }

plotLocalCases(area = 'South West', dat = dat)

```

## Milton Keynes 
```{r, echo=FALSE, fig.width=12, fig.height = 4 }

plotLocalCases(area = 'Milton Keynes', dat = dat)

```


## Hackney
```{r, echo=FALSE, fig.width=12, fig.height = 4 }

plotLocalCases(area = 'Hackney', dat = dat)

```


## Reading
```{r, echo=FALSE, fig.width=12, fig.height = 4 }

plotLocalCases(area = 'Reading', dat = dat)

```


```{r, echo=FALSE, fig.width=12, fig.height = 4 }

plotLocalCases(area = 'England', dat = dat)

```

```{r, echo=FALSE, fig.width=12, fig.height = 4 }

plotLocalCases(area = 'North West', dat = dat)

```

```{r, echo=FALSE, fig.width=12, fig.height = 4 }

plotLocalCases(area = 'North East', dat = dat)

```

```{r, echo=FALSE, fig.width=12, fig.height = 4 }

plotLocalCases(area = 'South East', dat = dat)

```


```{r, include=FALSE}

plotLocalRates <- function(area = 'Devon', dat, col = 'skyblue2', ...){
        # 'area' should be a string that names an area of interest
        
        localData <- dat[dat[,'Area.name']==area, ]
        
        da = seq(from = min(as.Date(localData[, 'Specimen.date'])), to = max(as.Date(localData[, 'Specimen.date'])), by = 7)

        
        par(las = 1, mar = c(2.5,4,4,1))
        plot(as.Date(localData[, 'Specimen.date']), localData[, 'Cumulative.lab.confirmed.cases.rate'],
             xlab = '', ylab = 'cases per 100 000 population', main = paste0('Cumulative lab confirmed cases rate in ', area),
             type = 'h', lwd = 5, col = col, bty = 'n', axes = FALSE, ...)
        grid()
        
        axis.Date(1,at=da,labels=format(da,"%d %b"),las=1)
        axis(2)
}

plotLocalRates(area = 'Devon', dat = dat)

```


```{r, fig.width=10, fig.height=4, include=FALSE, echo=FALSE}

par(las = 1)
plot(as.Date(devon[, 'Specimen.date']), devon[, 'Daily.lab.confirmed.cases'], type = 'n', axes = FALSE,
     xlab = 'Date', ylab = 'Daily lab confirmed cases', main = 'Daily lab confirmed Coronavirus cases in Devon & Exeter')
grid()
axis.Date(1, devon[, 'Specimen.date'])
axis(2)
lines(as.Date(devon[, 'Specimen.date']), devon[, 'Daily.lab.confirmed.cases'], type = 'o', pch = 20, col = 'black', lwd = 1.5)
lines(as.Date(exe[, 'Specimen.date']), exe[, 'Daily.lab.confirmed.cases'], type = 'o', pch = 20, col = 'blue', lwd = 1.5)
legend('topleft', legend = c('Devon', 'Exeter'), col = c('black', 'blue'), lty = 'solid', lwd = 1.5, pch = 20)
```



